---
title: "Seychelles Lit Review Results"
author: "**Pawel Waryszak**"
date: "Jan 28, 2021"
output:
  html_document: 
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
subtitle:  Deakin University, Blue Carbon Lab
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/BlueCarbon/Documents/00DeakinUni/R/BCL_R/Seychelles")
knitr::opts_chunk$set(dev='png')
```

# Results

```{r,echo=T, include=FALSE}
library(tidyverse)
library(readxl)
```


```{r,echo=T, include=FALSE}

#Looking at totla nuber of studies:
mg <- read_excel("SeychellesMangroveReview.xlsx", sheet = "Mangrove_BC_stock")
sg <- read_excel("SeychellesSeagrassReview.xlsx", sheet = "Seagrass_BC_stock")

mg_studies <- length(unique(mg$Reference)) #total n of studies on mangroves
sg_studies <- length(unique(sg$Reference)) #total n of studies on seagrass

mg_past10 <- filter(mg, Year > 2010)
mg_studies_past10 <- length(unique(mg_past10$Reference)) 

sg_past10 <- filter(sg, Year > 2010)
sg_studies_past10 <- length(unique(sg_past10$Reference)) 



#Looking at number of studies per region:
mg_studies_per_Country <- mg %>% group_by(Country) %>% summarize(mg_studies = n_distinct(Reference)) %>% arrange(desc(mg_studies)) %>% mutate(mg_tot=sum(mg_studies,na.rm=T))

sg_studies_per_Country <- sg %>% group_by(Country) %>% summarize(sg_studies = n_distinct(Reference))%>% arrange(desc(sg_studies)) %>% mutate(sg_tot=sum(sg_studies, na.rm=T))
sg_tot_value = mean(sg_studies_per_Country$sg_tot)#To replace NA that happen after left_join below:

mg_sg <- left_join(mg_studies_per_Country,sg_studies_per_Country , by = "Country", fill=0) %>% mutate(sg_studies = replace_na(sg_studies, 0)) %>% #Replacing Na with zero to avail summing
  mutate(sg_tot = replace_na(sg_tot, sg_tot_value)) %>% #replacing NA-s
  mutate (total = sg_tot + mg_tot)

top_mg <-  (mg_studies_per_Country[1,2]+ mg_studies_per_Country[2,2]) #total studies from top two countries
top_sg <-  (sg_studies_per_Country[1,2]+ sg_studies_per_Country[2,2]) #total studies from top two countries
top_two <- as.numeric(top_mg+top_sg)
tot  <-  as.numeric(mean(mg_sg$total)) #all sg + all mg


```


## Literature overview

Interest in blue carbon is growing rapidly in the SWIO research community, with 
`r round(((mg_studies_past10+sg_studies_past10)/(mg_studies+sg_studies)*100),1)`% of the studies identified published within the past 10 years (2011-2021, Fig. X). Folowing global trends, most of the research has largely focused on mangrove ecosystems and aboveground carbon stocks (see ‘Ecosystem’ and ‘Dataset’ below; Fig. Xa,b). Within the region, `r (mg_sg[1,1])` and `r mg_sg[2,1]` have taken the lead on blue carbon research, with `r round((top_two/tot*100),1)`% of the data being collected in these locations (See “Location’ below; Fig. Xa,b)

Despite the growing recognition on blue carbon stocks for climate change mitigation, few papers explicitly examined mangrove and seagrass carbon stocks. Many of the biomass or carbon measures reviewed were collected as complementary datasets to describe the environmental context/setting of field sites and experiments (e.g., ). Interestingly, many of the most complete and robust datasets were fueled by peer-reviewed publications arising from local Master and PhD theses (e.g., Njana2015, Githaiga2017, Juma2019), which suggests a rising interest in blue carbon research field.

## Ecosystem focus
From the `r tot` studies reviewed, `r round((mg_studies/tot*100), 1)`% were based on mangrove ecosystems, while `r round((sg_studies/tot*100), 1)`%  focused on seagrass meadows (Fig. Xa). Despite seagrass have higher extent in the WIO ( ha of seagrass vs ha of mangroves), mangrove research is likely favoured given (a) its applicability within REDD+ program; (b) its higher carbon storage and potential for carbon offsetting; (c) easier access to sites and sampling.  The need for additional datasets (eg. mangrove allometric equations and wood density) also increased the number of relevant publications within mangrove ecosystems.  

