---
title: "Seychelles Lit Review"
subtitle: "Overall Summary and Mangrove Results"
author: "**Pawel Waryszak**"
date: "Feb 11, 2021"
output:
  bookdown::word_document2:
    toc: yes
  #bookdown::html_document2
  #bookdown::pdf_document2:
keywords: "Seychelles Blue Carbon"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/BlueCarbon/Documents/00DeakinUni/R/BCL_R/Seychelles")
```

# Results

```{r,echo=T, include=FALSE}
library(tidyverse)
library(readxl)
```


```{r,echo=T,eval=T, include=FALSE}
#Looking at total nuber of studies:
mg <- read_excel("SeychellesMangroveReview.xlsx", sheet = "Mangrove_BC_stock")
sg <- read_excel("SeychellesSeagrassReview.xlsx", sheet = "Seagrass_BC_stock")

mg_studies <- length(unique(mg$Reference)) #total n of studies on mangroves
sg_studies <- length(unique(sg$Reference)) #total n of studies on seagrass

mg_past10 <- filter(mg, Year > 2010)
mg_studies_past10 <- length(unique(mg_past10$Reference)) 

sg_past10 <- filter(sg, Year > 2010)
sg_studies_past10 <- length(unique(sg_past10$Reference)) 

```



```{r,echo=T, include=FALSE}
#Looking at number of sg studies with AGC in Madagascar:
sg_AGC_studies_Madagascar <- sg %>%
  filter(is.na(AGC_MgCha)=="FALSE" | is.na(AGB_gDWm2)=="FALSE") %>%
  filter( Country == "Madagascar") %>%
  group_by(Reference)%>%
  summarize(sg_studies = n_distinct(Reference))

sg_AGC_studies_Madagascar_sum <- as.numeric(sum(sg_AGC_studies_Madagascar$sg_studies))

%>%
  arrange(desc(sg_studies))# %>%
  mutate(sg_tot=sum(sg_studies, na.rm=T))


#Looking at number of studies per region:
mg_studies_per_Country <- mg %>%
  group_by(Country) %>%
  summarize(mg_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_studies)) %>%
  mutate(mg_tot=sum(mg_studies,na.rm=T))

sg_studies_per_Country <- sg %>%
  group_by(Country) %>% 
  summarize(sg_studies = n_distinct(Reference))%>%
  arrange(desc(sg_studies)) %>%
  mutate(sg_tot=sum(sg_studies, na.rm=T))

sg_tot_value = mean(sg_studies_per_Country$sg_tot)#To replace NA that happen after left_join below:

mg_sg <- left_join(mg_studies_per_Country,sg_studies_per_Country , by = "Country", fill=0)%>%
  mutate(sg_studies = replace_na(sg_studies, 0)) %>%   #Replacing NA-s with zeros
  mutate(sg_tot = replace_na(sg_tot, sg_tot_value)) %>% #replacing NA-s with sg_tot_value
  mutate (total = sg_tot + mg_tot)

top_two_mg <-  (mg_studies_per_Country[1,2]+ mg_studies_per_Country[2,2]) #total studies from top two countries
top_two_sg <-  (sg_studies_per_Country[1,2]+ sg_studies_per_Country[2,2]) #total studies from top two countries

top_two <- as.numeric(top_two_mg + top_two_sg)
tot  <-  as.numeric(mean(mg_sg$total)) #all sg + all mg
sg_total <- as.numeric(mean(sg_studies_per_Country$sg_tot))
mg_total <- as.numeric(mean(mg_studies_per_Country$mg_tot))

```


## Literature overview

Interest in blue carbon is growing rapidly in the SWIO research community, with 
`r round(((mg_studies_past10+sg_studies_past10)/(mg_studies+sg_studies)*100),1)`% of the studies identified published within the past 10 years (2011-2021, Fig. X). Folowing global trends, most of the research has largely focused on mangrove ecosystems and aboveground carbon stocks (see ‘Ecosystem’ and ‘Dataset’ below; Fig. Xa,b). Within the region, `r (mg_sg[1,1])` and `r mg_sg[2,1]` have taken the lead on blue carbon research, with `r round((top_two/tot*100),1)`% of the data being collected in these locations (See “Location’ below; Fig. Xa,b).

Despite the growing recognition on blue carbon stocks for climate change mitigation, few papers explicitly examined mangrove and seagrass carbon stocks. Many of the biomass or carbon measures reviewed were collected as complementary datasets to describe the environmental context/setting of field sites and experiments (e.g., ). Interestingly, many of the most complete and robust datasets were fueled by peer-reviewed publications arising from local Master and PhD theses (e.g., Njana2015, Githaiga2017, Juma2019), which suggests a rising interest in blue carbon research field.

## Ecosystem focus

From the `r mg_studies + sg_studies` studies reviewed, `r round((mg_total/tot*100), 1)`% were based on mangrove ecosystems, while `r round((sg_total/tot*100), 1)`%  focused on seagrass meadows (Fig. Xa). Despite seagrass have higher extent in the WIO ( ha of seagrass vs ha of mangroves), mangrove research is likely favoured given (a) its applicability within REDD+ program; (b) its higher carbon storage and potential for carbon offsetting; (c) easier access to sites and sampling.  The need for additional datasets (eg. mangrove allometric equations and wood density) also increased the number of relevant publications within mangrove ecosystems.


## Datasets summary

```{r,echo=T,include=F}
##MANGROVES studies looking at AGB_MgDWha,BGB_MgDWha,AGC_MgCha,BGC_MgCha========
mg_plant_studies <- mg %>%
  filter(AGB_MgDWha>0 |BGB_MgDWha>0 | AGC_MgCha >0 | BGC_MgCha>0) %>%
  summarize(mg_plant_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_plant_studies))

mg_plant_studies_total <-sum(mg_plant_studies$mg_plant_studies,na.rm=T)

#Percent of mg plant studies with AboveGroundBiomass info:
mg_plant_studies_AGB <- mg %>%
  filter(AGB_MgDWha>0) %>%
  group_by(Country) %>%
  summarize(mg_plant_studies = n_distinct(Reference)) 

mg_plant_studies_AGB_total <- as.numeric(sum(mg_plant_studies_AGB$mg_plant_studies))
mg_AGB_MgDWha <- round((mg_plant_studies_AGB_total/mg_total*100),1)


#Percent of mg plant studies with BelowGroundBiomass info:
mg_plant_studies_BGB <- mg %>%
  filter(BGC_MgCha>0) %>%
  group_by(Country) %>%
  summarize(mg_plant_studies = n_distinct(Reference)) 

mg_plant_studies_BGB_total <- as.numeric(sum(mg_plant_studies_BGB$mg_plant_studies))
mg_BGC_MgCha <- round((mg_plant_studies_BGB_total/mg_total*100),1)


#SEAGRASS studies (looking at: AGB_gDWm2,AGC_MgCha,BGC_MgCh )===========
sg_plant_studies <- sg %>%
  filter(AGB_gDWm2>0 |BGB_gDWm2>0 | AGC_MgCha >0 | BGC_MgCha>0) %>%
  summarize(sg_plant_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_plant_studies))

sg_plant_studies_total <-sum(sg_plant_studies$sg_plant_studies,na.rm=T)
sg_AGB_MgDWha <- round((sg_plant_studies_total/sg_total*100),1)


#Percent of sg plant studies with AboveGroundBiomass info:
sg_plant_studies_AGB <- sg %>%
  filter(AGB_gDWm2>0) %>%
  summarize(sg_plant_studies = n_distinct(Reference)) 

sg_plant_studies_AGB_total <- as.numeric(sum(sg_plant_studies_AGB$sg_plant_studies))
sg_AGB_gDWm2 <- round((sg_plant_studies_AGB_total/sg_total*100),1)

#Percent of sg plant studies with BelowGroundBiomass info:
sg_plant_studies_BGB <- sg %>%
  filter(BGB_gDWm2>0) %>%
  summarize(sg_plant_studies = n_distinct(Reference)) 

sg_plant_studies_BGB_total <- as.numeric(sum(sg_plant_studies_BGB$sg_plant_studies))
sg_BGB_gDWm2 <- round((sg_plant_studies_BGB_total/sg_total*100),1)

#REF Examples:
#Aboveground biomass
mg_plant_studies_ref <- mg %>%
  filter(AGB_MgDWha > 0) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())

sg_plant_studies_ref <- sg %>%
  filter(AGB_gDWm2 > 0 ) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())

#Belowground biomasss:
mg_plant_studies_ref <- mg %>%
  filter(BGB_MgDWha > 0) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())%>% arrange(desc(ref_rows))

sg_plant_studies_ref <- sg %>%
  filter(BGB_gDWm2 > 0) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())%>% arrange(desc(ref_rows))


```


```{r,echo=T,include=F}
#Allometric Equations study count==========
#Load data sheet:
mg_eq <- read_excel("SeychellesMangroveReview.xlsx", sheet = "MG_Review")
sg_eq <- read_excel("SeychellesSeagrassReview.xlsx", sheet = "SG_Review")

#count PlantMetric for mangroves:
mg_papers_total <- as.numeric(length(unique(mg_eq$Paper)))
mg_plant_metrics <- as.numeric(length(mg_eq[mg_eq$PlantMetric == "Y", ]))
mg_eq_percent <- round((mg_plant_metrics/mg_papers_total*100),1)

#count PlantMetric for seagrasses:
sg_papers_total <- as.numeric(length(unique(sg_eq$Paper)))
sg_plant_metrics <- as.numeric(length(sg_eq[sg_eq$PlantMetric == "Y", ]))
sg_eq_percent <- round((sg_plant_metrics/sg_papers_total*100),1)


```

Most of the blue carbon literature in the WIO focuses on plant carbon stocks (`r sg_plant_studies_total + mg_plant_studies_total` studies; Fig. Xb); with `r mg_AGB_MgDWha`% of mangrove publications and `r sg_AGB_gDWm2`%  seagrass publications providing aboveground biomass (e.g., Cohen 2013 and Belshe 2018). Plant belowground biomass data were present in `r mg_BGC_MgCha` %  of mangrove and `r sg_BGB_gDWm2`% of seagrass studies (e.g.,Trettin 2016 and Gullstrom 2018). `r sg_eq_percent`% of seagrass and `r mg_eq_percent`% of mangrove studies are testing species- or site-specific allometric models (e.g., Abud 2017 and Arias-Ortiz 2020) to estimate plant biomass and/or related carbon stock. There is four mangrove studies that are reporting wood densities (e.g., Guillerot 2018). 

```{r,echo=T,include=F}
##MANGROVES studies on soil carbon stock (SOC_MgCha)========
mg_soil_studies <- mg %>%
  filter(SOC_MgCha>0) %>%
  summarize(mg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_soil_studies))

mg_soil_studies_total <-sum(mg_soil_studies$mg_soil_studies,na.rm=T)
mg_soil_OC_percent <- round((mg_soil_studies_total/mg_total*100),1)



#SEAGRASS studies on soil carbon stock ========
#List distinc seagrass studies:
sg_carbon_ref_studies <- sg %>%
  filter(is.na(AGC_MgCha)=="FALSE" | is.na(SOC_MgCha)=="FALSE") %>% #conatain no NA but numbers
  group_by(Reference) %>%
  summarize(carbon_studies = n_distinct(Reference)) %>%
  summarize(carbon_studies_sum = sum (carbon_studies))

sg_carbon_ref_studies_sum <- as.numeric(sg_carbon_ref_studies)
sg_carbon_ref_studies_percent <- round(sg_carbon_ref_studies_sum / sg_total *100, 1)

#SEAGRASS studies on soil carbon stock (SOC_MgCha) and biomass ========
#List distinc seagrass studies:
sg_soil_ref_studies <- sg %>%
  filter(SOC_MgCha>0) %>%
  group_by(Reference) %>%
  summarize(AV=mean(SOC_MgCha))

sg_soil_ref_studies_30cm <- sg %>%
  filter(DownToDepth_cm >30) %>%
  filter(SOC_MgCha>0) %>%
  group_by(Reference) %>%
  summarize(AV=mean(SOC_MgCha))

sg_30cm_soil_studies_refs <- as.character(sg_soil_ref_studies_30cm $ Reference)


#Number of distinct seagrass studies
sg_soil_studies <- sg %>%
  filter(SOC_MgCha>0) %>%
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_studies_total <-sum(sg_soil_studies$sg_soil_studies,na.rm=T)
sg_soil_OC_percent <- round((sg_soil_studies_total/sg_total*100),1)

#BGC or BGB count of seagrass studies:
sg_soil_biomass_studies <- sg %>%
  filter(BGC_MgCha>0  | BGB_gDWm2>0 ) %>%
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_BGB_BGC_studies_total <-sum( sg_soil_biomass_studies $ sg_soil_studies,na.rm=T)
sg_soil_BGB_BGC_percent <- round(( sg_soil_BGB_BGC_studies_total / sg_total*100),1)


#Shallow Core Depths========
shallow_depth_cm <- 10

#In mangroves
mg_soil_shallow <- mg %>%
  filter(DownToDepth_cm < shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(mg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_soil_studies))

mg_soil_shallow_percent <- as.numeric(round((mg_soil_shallow/mg_total*100),1))

#In seagrassses:
sg_soil_shallow <- sg %>%
  filter(DownToDepth_cm < shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_shallow_percent <- as.numeric(round((sg_soil_shallow/sg_total*100),1))


#Deep Core Depths========
#In mangroves:
mg_soil_deep <- mg %>%
  filter(DownToDepth_cm > shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(mg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_soil_studies))

mg_soil_deep_percent <- as.numeric(round((mg_soil_deep/mg_total*100),1))

#In seagrassses:
sg_soil_deep <- sg %>%
  filter(DownToDepth_cm > shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_deep_percent <- as.numeric(round((sg_soil_deep/sg_total*100),1))

```


```{r, echo=T, include=F}
#CN method description:
mg_CN_method <- mg %>%
  filter( SOC_Method !="NA") %>% #Remove blank rows
  group_by (SOC_Method) %>%
  summarize(mg_CN_types = n_distinct(Reference)) %>%
  arrange(desc(mg_CN_types))

#LOI method:
mg_LOI <- filter(mg_CN_method, SOC_Method=="LOI" )
mg_LOI_method_percent <- as.numeric(round((mg_LOI[1,2]/mg_total*100),1))

#CN_Analyser method:
mg_CN_Analyser <- filter(mg_CN_method, SOC_Method=="CN_Analyser" )
mg_CN_Analyser_method_percent <- as.numeric(round((mg_CN_Analyser[1,2]/mg_total*100),1))

#Walkley-Black Method (Pearson 2005):
mg_WBlack <- filter(mg_CN_method, SOC_Method=="Walkley-Black Method (Pearson 2005)" )
mg_WBlack_method_percent <- as.numeric(round((mg_WBlack[1,2]/mg_total*100),1)) 


#Age method description:
mg_age_method <- mg %>%
  filter(AgeDating_Method !="NA") %>% #Remove blank rows
  group_by (AgeDating_Method) %>%
  summarize(mg_age_types = n_distinct(Reference)) %>% #count distinct refs per SOC_Method
  arrange(desc(mg_age_types)) 

mg_age_total <-sum(mg_age_method$mg_age_types,na.rm=T)
mg_age_method_percent <- round((mg_age_total/mg_total*100),1)


```


Despite most of the blue carbon stocks are stored within the coastal sediments, only `r mg_soil_OC_percent`%  and `r sg_soil_OC_percent`% of the `r tot` studies included data on soil carbon stock in mangroves and seagrasses, respectively (Fig. Xb). `r mg_soil_shallow_percent`% of mangrove studies and `r sg_soil_shallow_percent`% of seagrass studies report carbon stocks values on shallow soil cores (shorter than `r as.character(shallow_depth_cm)` cm depth; e.g., Hemminga1994; Kristensen2008), while `r mg_soil_deep_percent`% and `r sg_soil_deep_percent`% provide carbon estimates along deeper soil profiles in mangroves and seagrasses, respectively (longer than 10 cm depth, e.g., Langat 2014; Gress 2017). Soil carbon content in the region is usually  calculated via the Walkley-Black Method (`r mg_WBlack_method_percent`%; e.g., Malgahaes 2019), loss-on-ignition (LOI; `r mg_LOI_method_percent`%; e.g., Lyimo 2002), or an elemental CN analyser (`r mg_CN_Analyser_method_percent`%; e.g., Middelburg 1996). Research that included soil carbon accretion rates was absent from seagrass studies and comprised small component of the mangrove literature (`r mg_age_method_percent`%). The age-dating in mangroves is generally focused on detecting Holocene sea-level changes using radiocarbon or Pb210 age dating on sediment cores (e.g., Punwong 2013, Woodoffre 2015). 


## Location
```{r, echo=T, include=F}
#Identified Seychelles Studies:=======
mg_Seychelles <- mg %>%
  filter( Country =="Seychelles") %>% 
  group_by(Location)%>%
  summarize(Seychelles_Studies = n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

mg_Seychelles_studies_total <-sum( mg_Seychelles $ Seychelles_Studies , na.rm=T)

#Mangrove Studies With C-data:
mg_Seychelles_data <- mg %>%
  filter( Country =="Seychelles") %>% 
  filter(AGB_MgDWha>0 |BGB_MgDWha>0 | AGC_MgCha >0 | BGC_MgCha>0 |SOC_MgCha>0) %>%
  group_by(Location)%>%
  summarize(Seychelles_Studies = n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

mg_Seychelles_studies_WithData <-sum( mg_Seychelles_data $ Seychelles_Studies , na.rm=T)


sg_Seychelles <- sg %>%
  filter( Country =="Seychelles") %>% 
  group_by(Location)%>%
  summarize(Seychelles_Studies= n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

sg_Seychelles_studies_total <-sum( sg_Seychelles $ Seychelles_Studies ,na.rm=T)

#Seagrass studies With C-data:
sg_Seychelles_data <- sg %>%
  filter( Country =="Seychelles") %>% 
  filter(AGB_gDWm2>0 |BGB_gDWm2>0 | AGC_MgCha >0 | BGC_MgCha>0) %>%
  group_by(Location)%>%
  summarize(Seychelles_Studies = n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

sg_Seychelles_studies_WithData <-sum( sg_Seychelles_data $ Seychelles_Studies , na.rm=T)


sg_mg_Seychelles_studies <- mg_Seychelles_studies_total + sg_Seychelles_studies_total
sg_mg_Seychelles_studies_WithData<-sg_Seychelles_studies_WithData+mg_Seychelles_studies_WithData 
```


Blue carbon research in the SWIO has been mainly conducted along the mainland East African coast (Figure 3), revealing a major geographical gap in the SIDS. Within the region, most studies have taken place in the coastal ecosystems of Kenya (mainly Gazi Bay), Tanzania (particularly Zanzibar Island and the Rufiji River Delta), and Mozambique (mainly Inhaca Island). Despite we identified `r sg_mg_Seychelles_studies` potential studies in Seychelles, only `r sg_mg_Seychelles_studies_WithData` included relevant datasets of coastal carbon stocks, emphasizing the need of further blue carbon research in the region.  

# Mangrove carbon stock summary 

## Trends in literature 

```{r,echo=T,include=F}
#C-Stock studies in mangroves==========
mg_stock <- mg %>%
  filter(AGB_MgDWha>0 |BGB_MgDWha>0 | AGC_MgCha >0 | BGC_MgCha>0 |SOC_MgCha>0) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(Stock_Studies))

mg_stock_studies <-sum( mg_stock $ Stock_Studies, na.rm=T)

#Aboveground C-stock studies:
mg_stock_aboveground <- mg %>%
  filter(AGB_MgDWha>0 | AGC_MgCha >0) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(Stock_Studies))

mg_stock_aboveground_studies <-sum( mg_stock_aboveground $ Stock_Studies, na.rm=T)

#Belowground C-stock studies:
mg_stock_belowground <- mg %>%
  filter(BGB_MgDWha>0 | BGC_MgCha>0) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(Stock_Studies))

mg_stock_belowground_studies <-sum( mg_stock_belowground $ Stock_Studies, na.rm=T)

```


A total of `r mg_stock_studies` studies reports mangrove carbon stocks from the WIO, including a recent review of African mangrove’s AGC, BGC and SOC by de Jong Cleyndert et al. (2020). The most of these studies aimed to quantify and describe the mangrove carbon stocks for a particular SWIO site or region. A number of the reional mangrove studies compared carbon stocks between certain ecosystem properties and stressors. For example: 

* Across canopy types or heights (Bandeira et al. 2009; Jones et al. 2015; Trettin et al. 2016) 

* From natural forests to those planted or reforested (e.g., Bosire et al. 2003; Tamooh et al. 2008; Kairo et al. 2009; Mutua et al. 2011) 

* Across different levels of human intervention (e.g., Kamau 2015, Bosire 2014, Msangameno 2017) 

From the identified `r mg_stock_studies` studies, `r mg_stock_aboveground_studies`% focused on ABG/AGC, while only `r mg_stock_belowground_studies`% included BGB/BGC datasets. The relatively few studies on mangrove BGB are likely due to the difficult and tedious excavation conditions required to sample the root and rhizome system of mangroves. 
 
Only `r mg_soil_studies_total` studies reported datasets on mangrove SOC stocks.   

Due to the lack of deep soils  .....

Many related mangrove studies that were not included on this review examined litter dynamics (e.g. Gwada & Kairo 2001; Fernando & Bandeira 2009), carbon fluxes (CH2, CH4; ), mangrove productivity (e.g., Githaiga2013) or stable isotopes (e.g., ).  

 
## Mangrove plant carbon

```{r ,fig.cap="Mangrove aboveground and belowground carbon stock by country in SWIO region  (tonnes ha^-1^ ,±SE)", echo=F,warning=F,message=F}

#Plot Mangrove Plant C-Stock by Country (mean +- SE):=========
#AboveGround Carbon:
mg_plant_above <- select (mg, Country,  BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha ) %>%
  #If "AGC_MgCha" present keep it  and if not we convert aboveground biomass to C by * 0.48
  mutate(Plant_C_Mgha = ifelse(AGC_MgCha > 0, AGC_MgCha, 0.48 * AGB_MgDWha)) %>%
  group_by(Country) %>%    #delete this line to get overall plant_C average
  summarise(AV=mean(Plant_C_Mgha, na.rm = T),
            SD=sd(Plant_C_Mgha, na.rm = T),
            N = length(Plant_C_Mgha),
            SE= SD / sqrt(N)) %>%
  mutate (Stock = "Aboveground")



#Belowground Carbon:
mg_plant_below <- select (mg, Country,  BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha ) %>%
  #If "BGC_MgCha" present iwe keep it  and if not we convert belowground biomass to C by * 0.39
  mutate(Plant_C_Mgha = ifelse(BGC_MgCha > 0, BGC_MgCha, 0.39 * BGB_MgDWha)) %>%
    group_by(Country) %>%    #delete this line to get overall plant_C average
    summarise(AV = mean(Plant_C_Mgha, na.rm = T),
            SD = sd(Plant_C_Mgha, na.rm = T),
            N  = length(Plant_C_Mgha),
            SE = SD / sqrt(N)) %>%
    mutate (Stock = "Belowground")

#Merge plant above and below carbo data together:
ab <- rbind (mg_plant_above, mg_plant_below)%>%
  #Turning belowground stock into negative values for plotting.
  mutate(AV= ifelse(Stock =="Aboveground",AV, AV*-1)) %>%
  na.omit()

#Draw a figure:
MyBreaks <- c(-300, -200,-100, -50, 0, 50, 100, 200,300)

plot_mangrove_plant <- ggplot(ab, aes(x=as.factor(Country), y=AV, fill = Stock))+
  geom_bar(position="identity", stat="identity")+
  geom_errorbar( aes(ymin= AV+SE, ymax = AV-SE), width=.4)+
  geom_hline(yintercept=0)+
  scale_y_continuous(breaks = MyBreaks,labels = abs(MyBreaks), limits = c(-300,300))+ #abs to remove negative values on y-axis below 0
  scale_fill_manual(values = c("#00FF00","#006633"))+
  xlab("Country")+ ylab(bquote("Plant carbon stock "  (Mg*~ha^-1)))+
  ggtitle("Mangroves")+
  theme_classic()+
  theme(axis.text.x=element_text(vjust=0.5,size=12),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=20),
        axis.title.x=element_text(size=20),
        legend.position = c(.2, .85),
        legend.text = element_text(size = 9),
        legend.title = element_text(face = "italic", size=10),
        plot.title = element_text(hjust = 0.5,lineheight=1.2, face="bold",size=20))

plot_mangrove_plant

#Boxplot of mangrove species:
box_sp <- select(mg,Species,AGC_MgCha)%>%
  na.omit() %>%
  ggplot( aes(x=as.factor(Species), y=AGC_MgCha, fill = Species))+
  geom_boxplot()+geom_jitter()


```


```{r,echo=T,include=F}
#Magrove plant C stock ranges======

#Ranges of aboveground carbon stocks:
min_AGC <- round(min(mg$AGC_MgCha, na.rm=T),1)
min_AGC_Country <- mg[which.min(mg$AGC_MgCha), "Country"]

max_AGC <- round(max(mg$AGC_MgCha, na.rm=T),1)
max_AGC_Country <- mg[which.max(mg$AGC_MgCha), "Country"]


#Ranges of belowground carbon stocks:
min_BGC <- round(min(mg$BGC_MgCha, na.rm=T),1)
max_BGC <- round(max(mg$BGC_MgCha, na.rm=T),1)

```



Mangrove AGC and BGC within the SWIO ranged between `r min_AGC`-`r max_AGC` tonnes C ha^-1^ and `r min_BGC`-`r max_BGC` tonnes C ha^-1^, respectively (Figure XYZ; Table S1). Although these values sit  within the global ranges reported for tropical mangroves (_____; REF), the wide AGC and BGC variability in the SWIO indicates mangrove stocks are largely influenced by the site conditions, species mix and forest structure (REF). Variability among mangrove species and locations may be further exacerbated by methodological or analytical differences across studies. For example, **check and compare 0.51400 t/ha in	Gang-Agatsiva1992 and 303.92157	t/ha in Alavaisha2016** (See also “Allometric Equations” below).


```{r ,fig.cap="Mangrove aboveground and belowground stock by species (tonnes ha^-1^ ,±SE)", echo=F,warning=F, message=F}

#Plot Mangrove Plant C-Stock by Species (mean +- SE):=========
#AboveGround Carbon:
mg_species_above <-select (mg, Species,BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha ) %>%
  #filter(Species !="Admixture")%>% #in case we want mono-stands only
    #If "AGC_MgCha" present keep it  and if not we convert aboveground biomass to C by * 0.48
  #*Most papers say "Stocks were determined from AGB and BGB using a carbon fraction of 0.48 for AGB and 0.39 for BGB
  #(Kauffman and Donato 2012; Abino et al. 2014).

  mutate(Plant_C_Mgha = ifelse(AGC_MgCha > 0, AGC_MgCha, 0.48 * AGB_MgDWha)) %>%
  group_by(Species) %>%    #delete this line to get overall plant_C average
  summarise(AV=mean(Plant_C_Mgha, na.rm = T),
            SD=sd(Plant_C_Mgha, na.rm = T),
            N = length(Plant_C_Mgha),
            SE= SD / sqrt(N)) %>%
  mutate (Stock = "Aboveground") %>%
    na.omit() 


#Belowground Carbon:
mg_species_below <-select(mg, Species,BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha) %>%
  #filter(Species !="Admixture")%>% #in case we want mono-stands only

  #If "AGC_MgCha" present keep it  and if not we convert aboveground biomass to C by * 0.48
  #*Most papers say "Stocks were determined from AGB and BGB using a carbon fraction of 0.48 for AGB and 0.39 for BGB
  #(Kauffman and Donato 2012; Abino et al. 2014).
  mutate(Plant_C_Mgha = ifelse(BGC_MgCha > 0, BGC_MgCha, 0.39 * BGB_MgDWha)) %>%
    group_by(Species) %>%    #delete this line to get overall plant_C average
    summarise(AV = mean(Plant_C_Mgha, na.rm = T),
            SD = sd(Plant_C_Mgha, na.rm = T),
            N  = length(Plant_C_Mgha),
            SE = SD / sqrt(N)) %>%
    mutate (Stock = "Belowground")%>%
      na.omit() 

#Merge plant above and below carbo data together:
ab_species <- rbind (mg_species_above, mg_species_below)%>%
  #Turning belowground stock into negative values for plotting.
  mutate(AV= ifelse(Stock =="Aboveground",AV, AV*-1)) %>%
  na.omit()


#Draw a figure:
MyBreaks_sp <- c( -200,-100, -50, 0, 50, 100, 200)

plot_mangrove_species <- ggplot(ab_species, aes(x=as.factor(Species), y=AV, fill = Stock))+
  geom_bar(position="identity", stat="identity")+
  geom_errorbar( aes(ymin= AV+SE, ymax = AV-SE), width=.4)+
  geom_hline(yintercept=0)+
  scale_y_continuous(breaks = MyBreaks_sp,labels = abs(MyBreaks_sp), limits = c(-300,300))+ #abs to remove negative values on y-axis below 0
  scale_fill_manual(values = c("#00FF00","#006633"))+
  xlab("Mangrove species")+ ylab(bquote("Plant carbon stock "  (Mg*~ha^-1)))+
  ggtitle("Mangroves")+
  theme_classic()+
  theme(axis.text.x=element_text(vjust=0.5,size=12),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=20),
        axis.title.x=element_text(size=20),
        legend.position = c(.2, .85),
        legend.text = element_text(size = 9),
        legend.title = element_text(face = "italic", size=10),
        plot.title = element_text(hjust = 0.5,lineheight=1.2, face="bold",size=20))

plot_mangrove_species

#Mean and SE Admixture Carbons:
mean_mg_carbon_mixed_above <- round(as.numeric(ab_species[ab_species$Species=="Admixture" & ab_species$Stock=="Aboveground","AV" ]),1)

SE_mg_carbon_mixed_above <- round(as.numeric(ab_species[ab_species$Species=="Admixture" & ab_species$Stock=="Aboveground","SE" ]),1)

```

```{r,echo=F,include=F,warning=F, message=F}
#C-stock range per  studies:========
#Compare methods for lowest and highest estimates of AGC between studies
mg_stock_range <- mg %>%
  filter(AGB_MgDWha>0 | AGC_MgCha >0) %>%
  group_by(AGC_MgCha,Reference) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(AGC_MgCha)) #0.51400	Gang-Agatsiva1992 and 303.92157	Alavaisha2016

#C-stock range per location:========
#Compare methods for lowest and highest estimates of AGC between sites (locations):
mg_stock_range_location <- mg %>%
  select(AGB_MgDWha, AGC_MgCha,Reference,Location,Country) %>%
  filter(AGB_MgDWha>0 | AGC_MgCha >0) %>%
  group_by(Reference) %>%
  mutate(Min_C = min(AGC_MgCha),
         Max_C = max(AGC_MgCha),
         C_Range = Max_C - Min_C )%>%
  #summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(C_Range)) #	Alavaisha2016


```



```{r, echo=F,warning=F, message=F,include=F}
#Min and Max Carbon stock within sites=======
Min_WithinSite <- as.numeric(mg_stock_range_location[1, "Min_C"])
Min_WithinSite_Location <- (mg_stock_range_location[1, "Location"])
Min_WithinSite_Country <- (mg_stock_range_location[1, "Country"])
High_WithinSiteVariation_Reference <- (mg_stock_range_location[1, "Reference"])


Max_WithinSite <- as.numeric(mg_stock_range_location[1, "Max_C"])

#C-stocks ranges in mixed mangrove forest:=======
AGC_mixed <- select(mg, AGC_MgCha, AGB_MgDWha, Species, Country, Location ) %>%
  filter(Species == "Admixture")%>% #=  mixed mangrove forest
  
#Turning biomass if no AGC into AGC (above ground carbon):
mutate(Plant_C_Mgha =ifelse(AGC_MgCha > 0, AGC_MgCha, 0.8 * AGB_MgDWha)) 

min_mixed_above_carbon <-  round(min(AGC_mixed$Plant_C_Mgha, na.rm=T),1)     
max_mixed_above_carbon <-  round(max(AGC_mixed$Plant_C_Mgha, na.rm=T),1)     

#C-stocks ranges within Species (monospecific stands):=======
AGC_species <- select(mg, AGC_MgCha, AGB_MgDWha, Species,Country, Location) %>%
filter(Species != "Admixture") %>% #Exclude mixed forest
  
#Turning biomass if no AGC into AGC (above ground carbon):
mutate(Plant_C_Mgha =ifelse(AGC_MgCha > 0, AGC_MgCha, 0.48 * AGB_MgDWha))  %>%
arrange(desc(Plant_C_Mgha))

#Cstock per species from:
min_sp_above_carbon <-  round(min(AGC_species$Plant_C_Mgha, na.rm=T),1)
min_sp_above_carbon_species  <- AGC_species[which.min(AGC_species$Plant_C_Mgha),"Species"]
min_sp_above_carbon_country <- AGC_species[which.min(AGC_species$Plant_C_Mgha),"Country"]

#Cstock per species to max of:
max_sp_above_carbon <-  round(max(AGC_species$Plant_C_Mgha, na.rm=T),1)
max_sp_above_carbon_species1  <- AGC_species[which.max(AGC_species$Plant_C_Mgha),"Species"]
max_sp_above_carbon_country1 <- AGC_species[which.max(AGC_species$Plant_C_Mgha),"Country"]

#C-stock of second max species:
AGC_species2 <- AGC_species [- which.max(AGC_species$Plant_C_Mgha), ]
max_sp_above_carbon_species2  <- AGC_species2[which.max(AGC_species2$Plant_C_Mgha),"Species"]
max_sp_above_carbon_country2 <- AGC_species2[which.max(AGC_species2$Plant_C_Mgha),"Country"]

#C-stock of third max species (all the same species, so I report only one):
AGC_species3 <- AGC_species2 [- which.max(AGC_species2$Plant_C_Mgha), ]
max_sp_above_carbon_species3  <- AGC_species3[which.max(AGC_species3$Plant_C_Mgha),"Species"]
max_sp_above_carbon_country3 <- AGC_species2[which.max(AGC_species3$Plant_C_Mgha),"Country"]



#Height, DBH, Density by mangrove species======
mg_height <- select(mg, Height_m,DBH_cm,TreeDensity_ha, Species)%>%
  group_by(Species) %>%
    summarise(AV_Height_m = mean(Height_m, na.rm = T),
            SD = sd(AV_Height_m, na.rm = T),
            N  = length(AV_Height_m),
            SE = SD / sqrt(N)) %>%
   arrange(desc(AV_Height_m))

mg_DBH_cm <- select(mg, Height_m,DBH_cm,TreeDensity_ha, Species)%>%
  group_by(Species) %>%
    summarise(AV_DBH_cm  = mean(Height_m, na.rm = T),
            SD = sd(AV_DBH_cm , na.rm = T),
            N  = length(AV_DBH_cm ),
            SE = SD / sqrt(N)) %>%
   arrange(desc(AV_DBH_cm ))

mg_TreeDensity_ha <- select(mg, Height_m,TreeDensity_ha,TreeDensity_ha, Species)%>%
  group_by(Species) %>%
    summarise(AV_TreeDensity_ha  = mean(Height_m, na.rm = T),
            SD = sd(AV_TreeDensity_ha , na.rm = T),
            N  = length(AV_TreeDensity_ha ),
            SE = SD / sqrt(N)) %>%
   arrange(desc(AV_TreeDensity_ha ))

max_sp_above_Height_m<-  round(max(mg$Height_m, na.rm=T),1)
max_mg_Height_m_sp1  <- mg[which.max(mg$Height_m),"Species"]

max_sp_above_DBH_cm <-  round(max(mg$DBH_cm, na.rm=T),1)
max_mg_height_sp1  <- mg[which.max(mg$DBH_cm),"Species"]


#Height, DBH, Density by seagrass species======
#Turn properties we are interested in to long format:
sg_properties_long <- select(sg, Height_m,Height_mDensity_m2,AGC_MgCha,BGC_MgCha,Reference) %>%
  gather(SG_Property, Value, -Reference)

#Use long format data  from mg_properties_long to computer percentages:
# sg_tot_value is the total number of studies on seagrass:

#Shoot %
sg_Shoot_Percent <- filter(sg, is.na(ShootDensity_m2)=="FALSE") %>%
  summarise(shoot_studies = n_distinct(Reference),
            shoot_studies_percent = round(shoot_studies/sg_tot_value*100),1)

sg_Shoot_Percent_Value <- as.numeric(sg_Shoot_Percent$shoot_studies_percent)# % shoots studies

#Height %:
sg_Height_m_Percent <- filter(sg, is.na(Height_m)=="FALSE") %>%
  summarise(Height_m_studies = n_distinct(Reference),
            Height_m_studies_percent = round(Height_m_studies/sg_tot_value*100),1)

sg_Height_m_Percent_Value <- as.numeric(sg_Height_m_Percent$Height_m_studies_percent)

#Most studied seagrass species:
sg_Species_n <- select(sg, Species,Reference) %>%
  group_by(Species) %>% 
  summarise(N_records= n()) %>%
  arrange(desc(N_records))

#Most studied mangrove species:
mg_Species_n <- select(mg, Species,Reference) %>%
  group_by(Species) %>% 
  summarise(N_records= n()) %>%
  arrange(desc(N_records))


```


AGC and BGC estimates varied considerably across the SWIO countries and sites. For example, the lowest mean AGC in mangroves of `r min_AGC` (tonnes ha^-1^)  was reported in `r min_AGC_Country` and the highest mean AGC of `r max_AGC` tonnes ha^-1^ in `r max_AGC_Country`. Variation across sites was evidenced in `r High_WithinSiteVariation_Reference`. Forests with a mix of mangrove species had a mean AGC of `r mean_mg_carbon_mixed_above` ±`r SE_mg_carbon_mixed_above`  tonnes C ha^-1^, ranging from `r min_mixed_above_carbon` to `r min_mixed_above_carbon` tonnes C ha^-1^. In terms of mangrove species, the highest AGC was recorded for *`r max_sp_above_carbon_species1`* stands  in `r max_sp_above_carbon_country1` (tonnes C ha^-1^; Fig. X; Table S1). The tallest height of `r max_sp_above_Height_m` m in the region  was reached by `r max_mg_Height_m_sp1`. The widest DBH of `r max_mg_height_sp1` was recorded in stand of *`r max_mg_height_sp1`*. The lowest AGC  of `r min_sp_above_carbon` (tonnes C ha^-1^) was recorded for *`r min_sp_above_carbon_species`*.


## Mangrove soil carbon

```{r, echo=F,warning=F, message=F}
#Plot Mangrove Soil C-Stock by Species and location (mean +- SE):=========

#Averages by Species:
mg_soil_C_Species <-select (mg, Species, SOC_MgCha, Bulkdensity_gDWmlorcm3, SOC_Percent,  DownToDepth_cm, SOM_Percent) %>%

  mutate(Soil_C_Mgha = ifelse(SOC_MgCha > 0, SOC_MgCha, 
                              Bulkdensity_gDWmlorcm3*SOC_Percent*DownToDepth_cm*100)) %>%
  
#mutate againg to include converstion from OM to OC:
#Soil organic matter generally contains approximately 56% organic carbon (Brady, 1990). The following #equation was therefore used to estimate percentage soil organic carbon (SOC) from total soil organic matter #(Brady, 1990): SOC% = SOM x 0.56


    mutate(Soil_C_Mgha2 = ifelse(Soil_C_Mgha > 0, Soil_C_Mgha, 
                              Bulkdensity_gDWmlorcm3*SOM_Percent*.56*DownToDepth_cm*100))  %>%

  group_by(Species) %>%    #delete this line to get overall Soil_C average
 
   summarise(AV= mean(Soil_C_Mgha2, na.rm = T),
            SD= sd(Soil_C_Mgha2, na.rm = T),
            N = length(Soil_C_Mgha2),
            SE= SD / sqrt(N)) %>%
  
  mutate(data = "Species") 

#Averages by Country:
mg_soil_C_Country <-select (mg, Country, SOC_MgCha,Bulkdensity_gDWmlorcm3, SOC_Percent,  DownToDepth_cm) %>%

  mutate(Soil_C_Mgha = ifelse(SOC_MgCha > 0, SOC_MgCha, 
                              Bulkdensity_gDWmlorcm3*SOC_Percent*DownToDepth_cm*100)) %>%
  
#mutate againg to include converstion from OM to OC:
#Soil organic matter generally contains approximately 56% organic carbon (Brady, 1990). The following #equation was therefore used to estimate percentage soil organic carbon (SOC) from total soil organic matter #(Brady, 1990): SOC% = SOM x 0.56

    mutate(Soil_C_Mgha2 = ifelse(Soil_C_Mgha > 0, Soil_C_Mgha, 
                              Bulkdensity_gDWmlorcm3*SOM_Percent*.56*DownToDepth_cm*100))  %>%

  group_by(Country) %>%    #delete this line to get overall Soil_C average
  
  summarise(AV= mean(Soil_C_Mgha2, na.rm = T),
            SD= sd(Soil_C_Mgha2, na.rm = T),
            N = length(Soil_C_Mgha2),
            SE= SD / sqrt(N))  %>%
    mutate(data = "Country") %>%
    rename(Species=Country) 

#Merge Soil above and below carbo data together:
ab_soil <- rbind (mg_soil_C_Country, mg_soil_C_Species) %>%na.omit()


plot_mangrove_soil <- ggplot(ab_soil, aes(x=Species, y=AV, fill=data))+
  geom_bar(position="identity", stat="identity")+
  geom_errorbar( aes(ymin= AV+SE, ymax = AV-SE), width=.4)+
  scale_fill_manual(values = c("#494928","#949447"))+
  facet_grid (.~data, scales = "free", space = "free")+
  xlab("")+
  ylab(bquote("Soil carbon stock "  (Mg*~ha^-1)))+
  ggtitle("Mangroves")+
  theme_bw()+
  theme(axis.text.x=element_text(vjust=0.5, size=12, angle=45),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        axis.title.x=element_text(size=15, angle=45),
        legend.position = "none",
        legend.text = element_text(size = 9),
        strip.text = element_text(size=15),
        legend.title = element_text(face = "italic", size=10),
        plot.title = element_text(hjust = 0.5,lineheight=1.2, face="bold",size=20))

plot_mangrove_soil

```





## Allometric equations.



```{r, echo=F,warning=F, message=F,results='asis'}

library(knitr)
mg_eq_table <- read_excel("SeychellesMangroveReview.xlsx", sheet = "AllometricEquations")

#Draw a table wtih Allometric equatons examples:
kable(mg_eq_table[,1:4], 
      caption="Table X. AGB and BGB allometric equations developed for mangrove species within the SWIO")


```

Most blue carbon studies worldwide use general equations published for mangrove forests with similar geomorphological characteristics (e.g., ). However, as recommended by **?????**,  several mangrove studies within the SWIO have developed species-specific or site-specific allometric equations (Table X).


# Seagrass Results

## TRENDS IN THE LITERATURE

The capacity of seagrass meadows to serve as carbon sinks has received limited scientific attention in the WIO. From the `r sg_total` relevant papers identified, only `r sg_carbon_ref_studies_percent`% of them directly examine seagrass carbon stocks (e.g., Githaiga 2017, Belshe et al. 2018, Gullström et al. 2018, Juma et al. 2020). The remaining studies report either seagrass biomass or carbon stocks as complimentary datasets to ecological or environmental studies (e.g., Mariani 1999, Ndaro and Ólafsson 1999, Vaïtilingon et al. 2003). As highlighted by Bandeira and Björk (2001), seagrass research in the region has largely emphasised on diversity, ecology and ecophysiology.

```{r, echo=F,warning=F, message=F}



```

Most blue carbon research has been conducted in Kenya and Tanzania thanks to the academic theses of Lyimo (2016), Dahl (2017), Githaiga (2017), Juma (2019), which have led to more than 10 publications in the topic (e.g., (Lyimo et al. 2006, Githaiga et al. 2016, 2017, Dahl et al. 2016, Gullström et al. 2018, Juma et al. 2020). Despite Madagascar holds most of the seagrass in the region, no studies have specifically focussed on carbon stocks and only `r sg_AGC_studies_Madagascar_sum` study included relevant information to estimate seagrass AGB and AGC (Vaïtilingon et al. 2003, Côté-Laurin et al. 2017). Our search only identified three relevant studies from Seychelles: all of them relating to seagrass biometrics and AGB (Aleem 1984, Kalugina-Gutnik et al. 1992, Ingram and Dawson 2001).

The majority of seagrass studies identified deal with plant _____. However only `r sg_soil_OC_percent`% of them directly estimate carbon stocks, with the rest limiting the datasets to seagrass biometrics or biomass. Unsurprisingly all of these studies measured seagrass aboveground stocks (AGB or AGC), but only `r sg_soil_BGB_BGC_percent`% included belowground datasets (BGB or BGC). As suggested by ______, this is likely due to the fact that most of the seagrass biomass/carbon is stored aboveground and considering the constrains of  sampling belowground rhizomes.  

Few studies have examined seagrass SOC stocks, demonstrating a major knowledge gap in the region. From the X studies identified, only three provide  SOC from sediment cores longer than >30 cm (Githaiga2017a; Belshe2018; Gullstrom2018). Several authors report seagrass SOC values from shallow sediment cores (<10 cm depth) when assessing the infauna from seagrass beds (e.g., Ndaro1999; Paula2001, Eklof2005).  


## Seagrass carbon 


```{r ,fig.cap="Seagrass aboveground and belowground carbon stock by country in SWIO region  (tonnes ha^-1^ ,±SE)", echo=F,warning=F,message=F}

#Plot Seagrass Plant C-Stock by Country (mean +- SE):=========
#AboveGround Carbon:
sg_plant_above <- select (sg, Country,  BGC_MgCha, AGC_MgCha, AGB_gDWm2, BGB_gDWm2)%>%
  #If "AGC_MgCha" present keep it  and if not we convert aboveground biomass to C by * 0.35
  mutate(Plant_C_Mgha = ifelse(AGC_MgCha > 0, AGC_MgCha, AGB_gDWm2*0.35/100 )) %>% #/100 to convert g/m2 to Mg/ha
  group_by(Country) %>%    #delete this line to get overall plant_C average
  summarise(AV=mean(Plant_C_Mgha, na.rm = T),
            SD=sd(Plant_C_Mgha, na.rm = T),
            N = length(Plant_C_Mgha),
            SE= SD / sqrt(N)) %>%
  mutate (Stock = "Aboveground")



#Belowground Carbon:
sg_plant_below <- select (sg, Country,  BGC_MgCha, AGC_MgCha,BGB_gDWm2) %>%
  # convert biomass to C by * 0.35 as per paper:
  #By Fourqurean2012 "Seagrass ecosystems as a globally significantcarbon stock"
  mutate(Plant_C_Mgha = ifelse(BGC_MgCha > 0, BGC_MgCha, BGB_gDWm2)) %>%
    group_by(Country) %>%    #delete this line to get overall plant_C average
    summarise(AV = mean(Plant_C_Mgha, na.rm = T),
            SD = sd(Plant_C_Mgha, na.rm = T),
            N  = length(Plant_C_Mgha),
            SE = SD / sqrt(N)) %>%
    mutate (Stock = "Belowground")

#Merge plant above and below carbo data together:
ab <- rbind (sg_plant_above, sg_plant_below)%>%
  #Turning belowground stock into negative values for plotting.
  mutate(AV= ifelse(Stock =="Aboveground",AV, AV*-1)) %>%
  na.omit()

#Draw a figure:
MyBreaks <- c(-7,-5,-3, -2,-1,  0, 1,2)

plot_Seagrass_plant <- ggplot(ab, aes(x=as.factor(Country), y=AV, fill = Stock))+
  geom_bar(position="identity", stat="identity")+
  geom_errorbar( aes(ymin= AV+SE, ymax = AV-SE), width=.4)+
  geom_hline(yintercept=0)+
  scale_y_continuous(breaks = MyBreaks,labels = abs(MyBreaks), limits = c(-7,2))+ #abs to remove negative values on y-axis below 0
  scale_fill_manual(values = c("#00FF00","#006633"))+
  xlab("Country")+ ylab(bquote("Plant carbon stock "  (Mg*~ha^-1)))+
  ggtitle("Seagrasses")+
  theme_classic()+
  theme(axis.text.x=element_text(vjust=0.5,size=12),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=20),
        axis.title.x=element_text(size=20),
        legend.position = c(.8, .2),
        legend.text = element_text(size = 9),
        legend.title = element_text(face = "italic", size=10),
        plot.title = element_text(hjust = 0.5,lineheight=1.2, face="bold",size=20))

plot_Seagrass_plant

```

```{r, echo=F,warning=F, message=F,include=F}
#Mean and SE Admixture Carbons:
mean_mg_carbon_mixed_above <- round(as.numeric(ab_species[ab_species$Species=="Admixture" & ab_species$Stock=="Aboveground","AV" ]),1)

SE_mg_carbon_mixed_above <- round(as.numeric(ab_species[ab_species$Species=="Admixture" & ab_species$Stock=="Aboveground","SE" ]),1)


```


Seagrass AGC and BGC within the SWIO had a mean of ___ ±___ ton C ha-1  (± SE) and ___ ±___  ton C ha-1, respectively (Figure 6; Table S1). These values are relatively __________ to the global medians reported by Fourqurean et al. (2012) of 0.26 ton C ha-1 for AGC and 0.54 ton C ha-1 for BGC. However, similar to global trends, WIO datasets reveal that approximately two-thirds of the living seagrass carbon is stored belowground. The seagrass meadows composed of mixed species had a AGC range of over ____ orders of magnitude (___ ±___ ton C ha-1), reflecting the importance of species composition. Within monospecific meadows, the highest seagrass AGC and BGC has been recorded for ________ followed by ________, _______ and _________; while the lowest was found for ______ in ______ (Figure 6). These pattern across species ???... 

 

 

Figure 6. Seagrass plant biomass/carbon stocks within the SWIO region. Aboveground and belowground biomass according to (a) species, and (b) location. 

 

SOIL CARBON  

 

The only available estimates were recently recorded in Tanzania (Belshe2018; Gullstrom2018) and Kenya (Githaiga2017a), where seagrass soil showed a mean carbon concentration of ___ ±___ % (± SE), and a SOC stock of ___ ±___  ton C ha-1 (X depth). These values are in line with ____________ , but relatively _______compared to the global median of 1.4% reported by Fourqurean et al (2012). 

 

Following global patterns, seagrass SOC stocks varied significantly across species and mixed meadows. ________ 

 

IDEAS 

Weird variability or random trends - could be related to tiny samples size OR super shallow cores.  

 

 

Fig. 7. Seagrass SOC stocks within the SWIO region. 


```{r, echo=F,warning=F, message=F,include=F}

```
