---
title: "Seychelles Lit Review Results"
author: "**Pawel Waryszak**"
date: "Jan 28, 2021"
output:
  bookdown::html_document2:
    toc: yes
  #bookdown::word_document2
  #bookdown::pdf_document2:
subtitle:  Deakin University, Blue Carbon Lab
keywords: "Seychelles Blue Carbon"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/BlueCarbon/Documents/00DeakinUni/R/BCL_R/Seychelles")
```

# Results

```{r,echo=T, include=FALSE}
library(tidyverse)
library(readxl)
```


```{r,echo=T, include=FALSE}
#Looking at total nuber of studies:
mg <- read_excel("SeychellesMangroveReview.xlsx", sheet = "Mangrove_BC_stock")
sg <- read_excel("SeychellesSeagrassReview.xlsx", sheet = "Seagrass_BC_stock")

mg_studies <- length(unique(mg$Reference)) #total n of studies on mangroves
sg_studies <- length(unique(sg$Reference)) #total n of studies on seagrass

mg_past10 <- filter(mg, Year > 2010)
mg_studies_past10 <- length(unique(mg_past10$Reference)) 

sg_past10 <- filter(sg, Year > 2010)
sg_studies_past10 <- length(unique(sg_past10$Reference)) 

```



```{r,echo=T, include=FALSE}
#Looking at number of studies per region:
mg_studies_per_Country <- mg %>%
  group_by(Country) %>%
  summarize(mg_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_studies)) %>%
  mutate(mg_tot=sum(mg_studies,na.rm=T))

sg_studies_per_Country <- sg %>%
  group_by(Country) %>% 
  summarize(sg_studies = n_distinct(Reference))%>%
  arrange(desc(sg_studies)) %>%
  mutate(sg_tot=sum(sg_studies, na.rm=T))

sg_tot_value = mean(sg_studies_per_Country$sg_tot)#To replace NA that happen after left_join below:

mg_sg <- left_join(mg_studies_per_Country,sg_studies_per_Country , by = "Country", fill=0)%>%
  mutate(sg_studies = replace_na(sg_studies, 0)) %>%   #Replacing NA-s with zeros
  mutate(sg_tot = replace_na(sg_tot, sg_tot_value)) %>% #replacing NA-s with sg_tot_value
  mutate (total = sg_tot + mg_tot)

top_two_mg <-  (mg_studies_per_Country[1,2]+ mg_studies_per_Country[2,2]) #total studies from top two countries
top_two_sg <-  (sg_studies_per_Country[1,2]+ sg_studies_per_Country[2,2]) #total studies from top two countries

top_two <- as.numeric(top_two_mg + top_two_sg)
tot  <-  as.numeric(mean(mg_sg$total)) #all sg + all mg
sg_total <- as.numeric(mean(sg_studies_per_Country$sg_tot))
mg_total <- as.numeric(mean(mg_studies_per_Country$mg_tot))

```


## Literature overview

Interest in blue carbon is growing rapidly in the SWIO research community, with 
`r round(((mg_studies_past10+sg_studies_past10)/(mg_studies+sg_studies)*100),1)`% of the studies identified published within the past 10 years (2011-2021, Fig. X). Folowing global trends, most of the research has largely focused on mangrove ecosystems and aboveground carbon stocks (see ‘Ecosystem’ and ‘Dataset’ below; Fig. Xa,b). Within the region, `r (mg_sg[1,1])` and `r mg_sg[2,1]` have taken the lead on blue carbon research, with `r round((top_two/tot*100),1)`% of the data being collected in these locations (See “Location’ below; Fig. Xa,b).

Despite the growing recognition on blue carbon stocks for climate change mitigation, few papers explicitly examined mangrove and seagrass carbon stocks. Many of the biomass or carbon measures reviewed were collected as complementary datasets to describe the environmental context/setting of field sites and experiments (e.g., ). Interestingly, many of the most complete and robust datasets were fueled by peer-reviewed publications arising from local Master and PhD theses (e.g., Njana2015, Githaiga2017, Juma2019), which suggests a rising interest in blue carbon research field.

## Ecosystem focus

From the `r tot` studies reviewed, `r round((mg_total/tot*100), 1)`% were based on mangrove ecosystems, while `r round((sg_total/tot*100), 1)`%  focused on seagrass meadows (Fig. Xa). Despite seagrass have higher extent in the WIO ( ha of seagrass vs ha of mangroves), mangrove research is likely favoured given (a) its applicability within REDD+ program; (b) its higher carbon storage and potential for carbon offsetting; (c) easier access to sites and sampling.  The need for additional datasets (eg. mangrove allometric equations and wood density) also increased the number of relevant publications within mangrove ecosystems.


## Datasets summary

```{r,echo=T,include=F}
##MANGROVES studies looking at AGB_MgDWha,BGB_MgDWha,AGC_MgCha,BGC_MgCha========
mg_plant_studies <- mg %>%
  filter(AGB_MgDWha>0 |BGB_MgDWha>0 | AGC_MgCha >0 | BGC_MgCha>0) %>%
  summarize(mg_plant_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_plant_studies))

mg_plant_studies_total <-sum(mg_plant_studies$mg_plant_studies,na.rm=T)

#Percent of mg plant studies with AboveGroundBiomass info:
mg_plant_studies_AGB <- mg %>%
  filter(AGB_MgDWha>0) %>%
  group_by(Country) %>%
  summarize(mg_plant_studies = n_distinct(Reference)) 

mg_plant_studies_AGB_total <- as.numeric(sum(mg_plant_studies_AGB$mg_plant_studies))
mg_AGB_MgDWha <- round((mg_plant_studies_AGB_total/mg_total*100),1)


#Percent of mg plant studies with BelowGroundBiomass info:
mg_plant_studies_BGB <- mg %>%
  filter(BGC_MgCha>0) %>%
  group_by(Country) %>%
  summarize(mg_plant_studies = n_distinct(Reference)) 

mg_plant_studies_BGB_total <- as.numeric(sum(mg_plant_studies_BGB$mg_plant_studies))
mg_BGC_MgCha <- round((mg_plant_studies_BGB_total/mg_total*100),1)


#SEAGRASS (looking at: AGB_gDWm2,AGC_MgCha,BGC_MgCh )===========
sg_plant_studies <- sg %>%
  filter(AGB_gDWm2>0 |BGB_gDWm2>0 | AGC_MgCha >0 | BGC_MgCha>0) %>%
  summarize(sg_plant_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_plant_studies))

sg_plant_studies_total <-sum(sg_plant_studies$sg_plant_studies,na.rm=T)
sg_AGB_MgDWha <- round((sg_plant_studies_total/sg_total*100),1)


#Percent of sg plant studies with AboveGroundBiomass info:
sg_plant_studies_AGB <- sg %>%
  filter(AGB_gDWm2>0) %>%
  summarize(sg_plant_studies = n_distinct(Reference)) 

sg_plant_studies_AGB_total <- as.numeric(sum(sg_plant_studies_AGB$sg_plant_studies))
sg_AGB_gDWm2 <- round((sg_plant_studies_AGB_total/sg_total*100),1)

#Percent of sg plant studies with BelowGroundBiomass info:
sg_plant_studies_BGB <- sg %>%
  filter(BGB_gDWm2>0) %>%
  summarize(sg_plant_studies = n_distinct(Reference)) 

sg_plant_studies_BGB_total <- as.numeric(sum(sg_plant_studies_BGB$sg_plant_studies))
sg_BGB_gDWm2 <- round((sg_plant_studies_BGB_total/sg_total*100),1)

#REF Examples:
#Aboveground biomass
mg_plant_studies_ref <- mg %>%
  filter(AGB_MgDWha > 0) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())

sg_plant_studies_ref <- sg %>%
  filter(AGB_gDWm2 > 0 ) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())

#Belowground biomasss:
mg_plant_studies_ref <- mg %>%
  filter(BGB_MgDWha > 0) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())%>% arrange(desc(ref_rows))

sg_plant_studies_ref <- sg %>%
  filter(BGB_gDWm2 > 0) %>%
  group_by(Reference) %>%
  summarize(ref_rows = n())%>% arrange(desc(ref_rows))


```


```{r,echo=T,include=F}
#Allometric Equations study count==========
#Load data sheet:
mg_eq <- read_excel("SeychellesMangroveReview.xlsx", sheet = "MG_Review")
sg_eq <- read_excel("SeychellesSeagrassReview.xlsx", sheet = "SG_Review")

#count PlantMetric for mangroves:
mg_papers_total <- as.numeric(length(unique(mg_eq$Paper)))
mg_plant_metrics <- as.numeric(length(mg_eq[mg_eq$PlantMetric == "Y", ]))
mg_eq_percent <- round((mg_plant_metrics/mg_papers_total*100),1)

#count PlantMetric for seagrasses:
sg_papers_total <- as.numeric(length(unique(sg_eq$Paper)))
sg_plant_metrics <- as.numeric(length(sg_eq[sg_eq$PlantMetric == "Y", ]))
sg_eq_percent <- round((sg_plant_metrics/sg_papers_total*100),1)


```

Most of the blue carbon literature in the WIO focuses on plant carbon stocks (`r sg_plant_studies_total + sg_plant_studies_total` studies; Fig. Xb); with `r mg_AGB_MgDWha`% of mangrove publications and `r sg_AGB_gDWm2`%  seagrass publications providing aboveground biomass (e.g., Cohen 2013 and Belshe 2018). Plant belowground biomass data were present in `r mg_BGC_MgCha` %  of mangrove and `r sg_BGB_gDWm2`% of seagrass studies (e.g.,Trettin 2016 and Gullstrom 2018). `r sg_eq_percent`% of seagrass and `r mg_eq_percent`% of mangrove studies are testing species- or site-specific allometric models (e.g., Abud 2017 and Arias-Ortiz 2020) to estimate plant biomass and/or related carbon stock. There is four mangrove studies that are reporting wood densities (e.g., Guillerot 2018). 

```{r,echo=T,include=F}
##MANGROVES studies on soil carbon stock (SOC_MgCha)========
mg_soil_studies <- mg %>%
  filter(SOC_MgCha>0) %>%
  summarize(mg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_soil_studies))

mg_soil_studies_total <-sum(mg_soil_studies$mg_soil_studies,na.rm=T)
mg_soil_OC_percent <- round((mg_soil_studies_total/mg_total*100),1)


#SEAGRASS studies on soil carbon stock (SOC_MgCha)========
sg_soil_studies <- sg %>%
  filter(SOC_MgCha>0) %>%
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_studies_total <-sum(sg_soil_studies$sg_soil_studies,na.rm=T)
sg_soil_OC_percent <- round((sg_soil_studies_total/sg_total*100),1)


#Shallow Core Depths========
shallow_depth_cm <- 10

#In mangroves
mg_soil_shallow <- mg %>%
  filter(DownToDepth_cm < shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(mg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_soil_studies))

mg_soil_shallow_percent <- as.numeric(round((mg_soil_shallow/mg_total*100),1))

#In seagrassses:
sg_soil_shallow <- sg %>%
  filter(DownToDepth_cm < shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_shallow_percent <- as.numeric(round((sg_soil_shallow/sg_total*100),1))


#Deep Core Depths========
#In mangroves:
mg_soil_deep <- mg %>%
  filter(DownToDepth_cm > shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(mg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(mg_soil_studies))

mg_soil_deep_percent <- as.numeric(round((mg_soil_deep/mg_total*100),1))

#In seagrassses:
sg_soil_deep <- sg %>%
  filter(DownToDepth_cm > shallow_depth_cm) %>%
  #group_by(Country)%>% # to see ref-s by country
  summarize(sg_soil_studies = n_distinct(Reference)) %>%
  arrange(desc(sg_soil_studies))

sg_soil_deep_percent <- as.numeric(round((sg_soil_deep/sg_total*100),1))

```


```{r, echo=T, include=F}
#CN method description:
mg_CN_method <- mg %>%
  filter( SOC_Method !="NA") %>% #Remove blank rows
  group_by (SOC_Method) %>%
  summarize(mg_CN_types = n_distinct(Reference)) %>%
  arrange(desc(mg_CN_types))

#LOI method:
mg_LOI <- filter(mg_CN_method, SOC_Method=="LOI" )
mg_LOI_method_percent <- as.numeric(round((mg_LOI[1,2]/mg_total*100),1))

#CN_Analyser method:
mg_CN_Analyser <- filter(mg_CN_method, SOC_Method=="CN_Analyser" )
mg_CN_Analyser_method_percent <- as.numeric(round((mg_CN_Analyser[1,2]/mg_total*100),1))

#Walkley-Black Method (Pearson 2005):
mg_WBlack <- filter(mg_CN_method, SOC_Method=="Walkley-Black Method (Pearson 2005)" )
mg_WBlack_method_percent <- as.numeric(round((mg_WBlack[1,2]/mg_total*100),1)) 


#Age method description:
mg_age_method <- mg %>%
  filter(AgeDating_Method !="NA") %>% #Remove blank rows
  group_by (AgeDating_Method) %>%
  summarize(mg_age_types = n_distinct(Reference)) %>% #count distinct refs per SOC_Method
  arrange(desc(mg_age_types)) 

mg_age_total <-sum(mg_age_method$mg_age_types,na.rm=T)
mg_age_method_percent <- round((mg_age_total/mg_total*100),1)


```


Despite most of the blue carbon stocks are stored within the coastal sediments, only `r mg_soil_OC_percent`%  and `r sg_soil_OC_percent`% of the `r tot` studies included data on soil carbon stock in mangroves and seagrasses, respectively (Fig. Xb). `r mg_soil_shallow_percent`% of mangrove studies and `r sg_soil_shallow_percent`% of seagrass studies report carbon stocks values on shallow soil cores (shorter than `r as.character(shallow_depth_cm)` cm depth; e.g., Hemminga1994; Kristensen2008), while `r mg_soil_deep_percent`% and `r sg_soil_deep_percent`% provide carbon estimates along deeper soil profiles in mangroves and seagrasses, respectively (longer than 10 cm depth, e.g., Langat 2014; Gress 2017). Soil carbon content in the region is usually  calculated via the Walkley-Black Method (`r mg_WBlack_method_percent`%; e.g., Malgahaes 2019), loss-on-ignition (LOI; `r mg_LOI_method_percent`%; e.g., Lyimo 2002), or an elemental CN analyser (`r mg_CN_Analyser_method_percent`%; e.g., Middelburg 1996). Research that included soil carbon accretion rates was absent from seagrass studies and comprised small component of the mangrove literature (`r mg_age_method_percent`%). The age-dating in mangroves is generally focused on detecting Holocene sea-level changes using radiocarbon or Pb210 age dating on sediment cores (e.g., Punwong 2013, Woodoffre 2015). 


## Location
```{r, echo=T, include=F}
#Identified Seychelles Studies:=======
mg_Seychelles <- mg %>%
  filter( Country =="Seychelles") %>% 
  group_by(Location)%>%
  summarize(Seychelles_Studies = n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

mg_Seychelles_studies_total <-sum( mg_Seychelles $ Seychelles_Studies , na.rm=T)

#Mangrove Studies With C-data:
mg_Seychelles_data <- mg %>%
  filter( Country =="Seychelles") %>% 
  filter(AGB_MgDWha>0 |BGB_MgDWha>0 | AGC_MgCha >0 | BGC_MgCha>0 |SOC_MgCha>0) %>%
  group_by(Location)%>%
  summarize(Seychelles_Studies = n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

mg_Seychelles_studies_WithData <-sum( mg_Seychelles_data $ Seychelles_Studies , na.rm=T)


sg_Seychelles <- sg %>%
  filter( Country =="Seychelles") %>% 
  group_by(Location)%>%
  summarize(Seychelles_Studies= n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

sg_Seychelles_studies_total <-sum( sg_Seychelles $ Seychelles_Studies ,na.rm=T)

#Seagrass studies With C-data:
sg_Seychelles_data <- sg %>%
  filter( Country =="Seychelles") %>% 
  filter(AGB_gDWm2>0 |BGB_gDWm2>0 | AGC_MgCha >0 | BGC_MgCha>0) %>%
  group_by(Location)%>%
  summarize(Seychelles_Studies = n_distinct(Reference)) %>%
  arrange(desc(Seychelles_Studies))

sg_Seychelles_studies_WithData <-sum( sg_Seychelles_data $ Seychelles_Studies , na.rm=T)


sg_mg_Seychelles_studies <- mg_Seychelles_studies_total + sg_Seychelles_studies_total
sg_mg_Seychelles_studies_WithData<-sg_Seychelles_studies_WithData+mg_Seychelles_studies_WithData 
```


Blue carbon research in the SWIO has been mainly conducted along the mainland East African coast (Figure 3), revealing a major geographical gap in the SIDS. Within the region, most studies have taken place in the coastal ecosystems of Kenya (mainly Gazi Bay), Tanzania (particularly Zanzibar Island and the Rufiji River Delta), and Mozambique (mainly Inhaca Island). Despite we identified `r sg_mg_Seychelles_studies` potential studies in Seychelles, only `r sg_mg_Seychelles_studies_WithData` included relevant datasets of coastal carbon stocks, emphasizing the need of further blue carbon research in the region.  

# Mangrove carbon stock summary 

## Trends in literature 

```{r,echo=T,include=F}
#C-Stock studies in mangroves==========
mg_stock <- mg %>%
  filter(AGB_MgDWha>0 |BGB_MgDWha>0 | AGC_MgCha >0 | BGC_MgCha>0 |SOC_MgCha>0) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(Stock_Studies))

mg_stock_studies <-sum( mg_stock $ Stock_Studies, na.rm=T)

#Aboveground C-stock studies:
mg_stock_aboveground <- mg %>%
  filter(AGB_MgDWha>0 | AGC_MgCha >0) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(Stock_Studies))

mg_stock_aboveground_studies <-sum( mg_stock_aboveground $ Stock_Studies, na.rm=T)

#Belowground C-stock studies:
mg_stock_belowground <- mg %>%
  filter(BGB_MgDWha>0 | BGC_MgCha>0) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(Stock_Studies))

mg_stock_belowground_studies <-sum( mg_stock_belowground $ Stock_Studies, na.rm=T)

```


A total of `r mg_stock_studies` studies reports mangrove carbon stocks from the WIO, including a recent review of African mangrove’s AGC, BGC and SOC by de Jong Cleyndert et al. (2020). The most of these studies aimed to quantify and describe the mangrove carbon stocks for a particular SWIO site or region. A number of the reional mangrove studies compared carbon stocks between certain ecosystem properties and stressors. For example: 

* Across canopy types or heights (Bandeira et al. 2009; Jones et al. 2015; Trettin et al. 2016) 

* From natural forests to those planted or reforested (e.g., Bosire et al. 2003; Tamooh et al. 2008; Kairo et al. 2009; Mutua et al. 2011) 

* Across different levels of human intervention (e.g., Kamau 2015, Bosire 2014, Msangameno 2017) 

From the identified `r mg_stock_studies` studies, `r mg_stock_aboveground_studies`% focused on ABG/AGC, while only `r mg_stock_belowground_studies`% included BGB/BGC datasets. The relatively few studies on mangrove BGB are likely due to the difficult and tedious excavation conditions required to sample the root and rhizome system of mangroves. 
 
Only `r mg_soil_studies_total` studies reported datasets on mangrove SOC stocks.   

Due to the lack of deep soils  .....

Many related mangrove studies that were not included on this review examined litter dynamics (e.g. Gwada & Kairo 2001; Fernando & Bandeira 2009), carbon fluxes (CH2, CH4; ), mangrove productivity (e.g., Githaiga2013) or stable isotopes (e.g., ).  

 
## Mangrove C stock

```{r ,fig.cap="Mangrove aboveground and belowground Stock (tonnes ha^-1^ ,+SE)", echo=F,warning=F}

#Plot Mangrove Plant C-Stock by Country (mean +- SE):=========
#AboveGround Carbon:
mg_plant_above <- select (mg, Country,  BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha ) %>%
  #If "AGC_MgCha" present keep it  and if not we convert aboveground biomass to C by * 0.464
  mutate(Plant_C_Mgha = ifelse(AGC_MgCha > 0, AGC_MgCha, 0.464 * AGB_MgDWha)) %>%
  group_by(Country) %>%    #delete this line to get overall plant_C average
  summarise(AV=mean(Plant_C_Mgha, na.rm = T),
            SD=sd(Plant_C_Mgha, na.rm = T),
            N = length(Plant_C_Mgha),
            SE= SD / sqrt(N)) %>%
  mutate (Stock = "Aboveground")


#Belowground Carbon:
mg_plant_below <- select (mg, Country,  BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha ) %>%
  #If "BGC_MgCha" present iwe keep it  and if not we convert belowground biomass to C by * 0.464
  mutate(Plant_C_Mgha = ifelse(BGC_MgCha > 0, BGC_MgCha, 0.464 * BGB_MgDWha)) %>%
    group_by(Country) %>%    #delete this line to get overall plant_C average
    summarise(AV = mean(Plant_C_Mgha, na.rm = T),
            SD = sd(Plant_C_Mgha, na.rm = T),
            N  = length(Plant_C_Mgha),
            SE = SD / sqrt(N)) %>%
    mutate (Stock = "Belowground")

#Merge plant above and below carbo data together:
ab <- rbind (mg_plant_above, mg_plant_below)%>%
  #Turning belowground stock into negative values for plotting.
  mutate(AV= ifelse(Stock =="Aboveground",AV, AV*-1)) %>%
  na.omit()

#Draw a figure:
MyBreaks <- c(-300, -200,-100, -50, 0, 50, 100, 200,300)

plot_mangrove_plant <- ggplot(ab, aes(x=as.factor(Country), y=AV, fill = Stock))+
  geom_bar(position="identity", stat="identity")+
  geom_errorbar( aes(ymin= AV+SE, ymax = AV-SE), width=.4)+
  geom_hline(yintercept=0)+
  scale_y_continuous(breaks = MyBreaks,labels = abs(MyBreaks), limits = c(-300,300))+ #abs to remove negative values on y-axis below 0
  scale_fill_manual(values = c("#00FF00","#006633"))+
  xlab("Country")+ ylab(bquote("Plant carbon stock " (Mg*~ha^-1)))+
  ggtitle("Mangroves")+
  theme_classic()+
  theme(axis.text.x=element_text(vjust=0.5,size=12),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=20),
        legend.position = c(.2, .85),
        legend.text = element_text(size = 9),
        legend.title = element_text(face = "italic", size=10),
        plot.title = element_text(hjust = 0.5,lineheight=1.2, face="bold",size=20))

plot_mangrove_plant



```


```{r,echo=T,include=F}
#Plant C stock ranges======

#Ranges of aboveground carbon stocks:
min_AGC <- round(min(mg$AGC_MgCha, na.rm=T),1)
max_AGC <- round(max(mg$AGC_MgCha, na.rm=T),1)


#Ranges of belowground carbon stocks:
min_BGC <- round(min(mg$BGC_MgCha, na.rm=T),1)
max_BGC <- round(max(mg$BGC_MgCha, na.rm=T),1)

#Range og stocks per studies:========
#Compare methods for lowest and highest estimates of AGB
mg_stock_range <- mg %>%
  filter(AGB_MgDWha>0 | AGC_MgCha >0) %>%
  group_by(AGC_MgCha,Reference) %>%
  summarize(Stock_Studies = n_distinct(Reference)) %>%
  arrange(desc(AGC_MgCha)) #0.51400	Gang-Agatsiva1992 and 303.92157	Alavaisha2016


```


Mangrove AGC and BGC within the SWIO ranged between `r min_AGC`-`r max_AGC` tonnes C ha^-1^ and `r min_BGC`-`r max_BGC` tonnes C ha^-1^, respectively (Figure XYZ; Table S1). Although these values sit  within the global ranges reported for tropical mangroves (_____; REF), the wide AGC and BGC variability in the SWIO indicates mangrove stocks are largely influenced by the site conditions, species mix and forest structure (REF). Variability among mangrove species and locations may be further exacerbated by methodological or analytical differences across studies. For example, **check and compare 0.51400 t/ha in	Gang-Agatsiva1992 and 303.92157	t/ha in Alavaisha2016** (See also “Allometric Equations” below).


```{r ,fig.cap="Mangrove aboveground and belowground stock by species (tonnes ha^-1^ ,+SE)", echo=F,warning=F}

#Plot Mangrove Plant C-Stock by Species (mean +- SE):=========
#AboveGround Carbon:
mg_species_above <-select(mg,Species,BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha ) %>%
  filter(Species !="Admixture")%>%
  #we convert aboveground biomass to C by * 0.464 if not reported:
  mutate(Plant_C_Mgha = ifelse(AGC_MgCha > 0, AGC_MgCha, 0.464 * AGB_MgDWha)) %>%
  group_by(Species) %>%    #delete this line to get overall plant_C average
  summarise(AV=mean(Plant_C_Mgha, na.rm = T),
            SD=sd(Plant_C_Mgha, na.rm = T),
            N = length(Plant_C_Mgha),
            SE= SD / sqrt(N)) %>%
  mutate (Stock = "Aboveground") %>%
    na.omit() 


#Belowground Carbon:
mg_species_below <-select(mg, Species,BGC_MgCha, AGC_MgCha, AGB_MgDWha, BGB_MgDWha) %>%
    filter(Species !="Admixture")%>%

  #we convert aboveground biomass to C by * 0.464 if not reported:
  mutate(Plant_C_Mgha = ifelse(BGC_MgCha > 0, BGC_MgCha, 0.464 * BGB_MgDWha)) %>%
    group_by(Species) %>%    #delete this line to get overall plant_C average
    summarise(AV = mean(Plant_C_Mgha, na.rm = T),
            SD = sd(Plant_C_Mgha, na.rm = T),
            N  = length(Plant_C_Mgha),
            SE = SD / sqrt(N)) %>%
    mutate (Stock = "Belowground")%>%
      na.omit() 

#Merge plant above and below carbo data together:
ab_species <- rbind (mg_species_above, mg_species_below)%>%
  #Turning belowground stock into negative values for plotting.
  mutate(AV= ifelse(Stock =="Aboveground",AV, AV*-1)) %>%
  na.omit()

#Draw a figure:
MyBreaks_sp <- c( -200,-100, -50, 0, 50, 100, 200)

plot_mangrove_species <- ggplot(ab_species, aes(x=as.factor(Species), y=AV, fill = Stock))+
  geom_bar(position="identity", stat="identity")+
  geom_errorbar( aes(ymin= AV+SE, ymax = AV-SE), width=.4)+
  geom_hline(yintercept=0)+
  scale_y_continuous(breaks = MyBreaks_sp,labels = abs(MyBreaks_sp), limits = c(-300,300))+ #abs to remove negative values on y-axis below 0
  scale_fill_manual(values = c("#00FF00","#006633"))+
  xlab("Mangrove species")+ ylab(bquote("Plant carbon stock " (Mg*~ha^-1)))+
  ggtitle("Mangrove species")+
  theme_classic()+
  theme(axis.text.x=element_text(vjust=0.5,size=12),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=20),
        legend.position = c(.2, .85),
        legend.text = element_text(size = 9),
        legend.title = element_text(face = "italic", size=10),
        plot.title = element_text(hjust = 0.5,lineheight=1.2, face="bold",size=20))

plot_mangrove_species


```




AGC and BGC estimates varied considerably across the SWIO countries and sites. For example, _______________. Variation across sites was evidenced in _______. Forests with a mix of mangrove species had a mean AGC of ___ ±___ ton C ha-1, ranging from ____ to _____ ton C ha-1. In terms of mangrove species, the highest AGC was recorded for monospecific stands of _______ in ________ (ton C ha-1), _______ in __________ (ton C ha-1) and __________ in __________ (ton C ha-1; Fig. X; Table S1). These three species most often store the highest AGC stocks in the region given they tend to reach the tallest heights (> X m) and hold the widest DBH (> X cm). The lowest AGC was recorded for the short mangrove ________ (ton C ha-1) who rarely exceeds X m height or X cm DBH.  
